---
title: "Integrate DC datasets"
author: "Massimo Andreatta & Santiago Carmona"
#output: html_notebook
---

```{r}
renv::restore()
```

Libraries
```{r}
#remotes::install_github('carmonalab/STACAS')
#remotes::install_github('carmonalab/scIntegrationMetrics')

library(ggplot2)
library(Seurat)
library(tidyr)
library(SeuratObject)
library(Matrix)
library(patchwork)
library(remotes)
library(UCell)
library(STACAS)
library(scIntegrationMetrics)
```

Read in objects for uncorrected, unsupervised and semi-sup integration
```{r}
merged <- readRDS("aux/DC_unintegrated.rds")

ref.unsup <- readRDS("aux/DC_reference_wTcells.rds")

ref.sup <- readRDS("aux/DC_reference_2nd_wmeta.rds")
```


```{r fig.height=2, fig.width=5}
a <- DimPlot(merged, group.by = "common.name") + NoLegend() + theme(aspect.ratio = 1)
b <- DimPlot(ref.unsup, group.by = "common.name") + NoLegend() + theme(aspect.ratio = 1)
c <- DimPlot(ref.sup, group.by = "common.name") + theme(aspect.ratio = 1)
a | b | c
```

Subset on same cells (doublets and contaminants were removed on final iteration)
```{r}
cells <- colnames(ref.sup)
merged <- subset(merged, cells=cells)
ref.unsup <- subset(ref.unsup, cells=cells)
```

Calculate integration metrics based on annotations by Gerhard et al.
```{r}
#Integration metrics on uncorrected data
integrationMetrics <- list()

metricsLabels <- c("cDC1","cDC2/MoDC","DC3","pDC")  #cannot evaluate AS-DC, as they are only annotated in one dataset

integrationMetrics[["Uncorrected"]] <- getIntegrationMetrics(object=merged,
                                                      meta.label = "common.name",
                                                      meta.batch = "Study",
                                                      metricsLabels = metricsLabels)

integrationMetrics[["Unsupervised"]] <- getIntegrationMetrics(object=ref.unsup,
                                                      meta.label = "common.name",
                                                      meta.batch = "Study",
                                                      metricsLabels = metricsLabels)

integrationMetrics[["Semi-supervised"]] <- getIntegrationMetrics(object=ref.sup,
                                                      meta.label = "common.name",
                                                      meta.batch = "Study",
                                                      metricsLabels = metricsLabels)
```

Combine integration metrics
```{r fig.height=3}
integrationMetricsSummary <- data.frame(unlist(integrationMetrics)) %>%
  tibble::rownames_to_column() %>%
  dplyr::rename(value=unlist.integrationMetrics.) %>%
  separate(rowname, c("Method","Metric"), sep="\\.")


integrationMetricsSummary$Method <- factor(integrationMetricsSummary$Method,
                                           levels=c("Uncorrected","Unsupervised","Semi-supervised"))

ggplot(integrationMetricsSummary, aes(x=Method, y=value, fill=Method)) + geom_bar(stat="identity") + 
    theme_bw()  +
    theme(legend.position="none",
          axis.text.x=element_text(angle = 45, vjust = 1, hjust=1)) +
    xlab("Method") + facet_wrap(~Metric, scales = "free") +
    scale_fill_brewer(palette="Set1")
```


Scatterplot of two performance metrics
```{r fig.width=4, fig.height=1.5}
palette <- brewer.pal(n=length(unique(integrationMetricsSummary$Method)), name = "Set1")

a <- integrationMetricsSummary %>% dplyr::filter(Metric %in% c("celltype_ASW","CiLISI")) %>%
  pivot_wider(names_from = Metric, values_from = value ) %>%
  ggplot(aes(x=CiLISI, y=`celltype_ASW`, label=Method)) +
  geom_point(aes(color=Method), size=2) + geom_text(hjust=0.1, vjust=0) + 
  scale_color_manual(values = palette) + theme_light()

b <- integrationMetricsSummary %>% dplyr::filter(Metric %in% c("norm_cLISI","CiLISI")) %>%
  pivot_wider(names_from = Metric, values_from = value ) %>%
  ggplot(aes(x=CiLISI, y=`norm_cLISI`, label=Method)) +
  geom_point(aes(color=Method), size=2) + geom_text(hjust=0.1, vjust=0) + 
  scale_color_manual(values = palette) + theme_light()

a | b

ggsave("plots/DC_metrics_scatterplot.pdf", height=3, width=10)
```

# Metric on expert-curated labels
```{r}
not.doublets <- colnames(ref.unsup)[ref.unsup$functional.cluster != "Tcell_doublets"]

merged <- subset(merged, cells=not.doublets)
ref.unsup <- subset(ref.unsup, cells=not.doublets)
ref.sup <- subset(ref.sup, cells=not.doublets)

labels <- as.character(ref.unsup$functional.cluster)
names(labels) <- colnames(ref.unsup)

merged$expert.labels <- NA
merged$expert.labels[names(labels)] <- labels

ref.unsup$expert.labels <- NA
ref.unsup$expert.labels[names(labels)] <- labels

ref.sup$expert.labels <- NA
ref.sup$expert.labels[names(labels)] <- labels

```


```{r}
#Integration metrics on uncorrected data
integrationMetrics <- list()

metricsLabels <- NULL  #use all clusters

integrationMetrics[["Uncorrected"]] <- getIntegrationMetrics(object=merged,
                                                      meta.label = "expert.labels",
                                                      meta.batch = "Study",
                                                      metricsLabels = metricsLabels)

integrationMetrics[["Unsupervised"]] <- getIntegrationMetrics(object=ref.unsup,
                                                      meta.label = "expert.labels",
                                                      meta.batch = "Study",
                                                      metricsLabels = metricsLabels)

integrationMetrics[["Semi-supervised"]] <- getIntegrationMetrics(object=ref.sup,
                                                      meta.label = "expert.labels",
                                                      meta.batch = "Study",
                                                      metricsLabels = metricsLabels)

integrationMetricsSummary <- data.frame(unlist(integrationMetrics)) %>%
  tibble::rownames_to_column() %>%
  dplyr::rename(value=unlist.integrationMetrics.) %>%
  separate(rowname, c("Method","Metric"), sep="\\.")


integrationMetricsSummary$Method <- factor(integrationMetricsSummary$Method,
                                           levels=c("Uncorrected","Unsupervised","Semi-supervised"))

ggplot(integrationMetricsSummary, aes(x=Method, y=value, fill=Method)) + geom_bar(stat="identity") + 
    theme_bw()  +
    theme(legend.position="none",
          axis.text.x=element_text(angle = 45, vjust = 1, hjust=1)) +
    xlab("Method") + facet_wrap(~Metric, scales = "free") +
    scale_fill_brewer(palette="Set1")
```
Scatterplot of two performance metrics
```{r fig.width=4, fig.height=1.5}
palette <- brewer.pal(n=length(unique(integrationMetricsSummary$Method)), name = "Set1")

a <- integrationMetricsSummary %>% dplyr::filter(Metric %in% c("celltype_ASW","CiLISI")) %>%
  pivot_wider(names_from = Metric, values_from = value ) %>%
  ggplot(aes(x=CiLISI, y=`celltype_ASW`, label=Method)) +
  geom_point(aes(color=Method), size=2) + geom_text(hjust=0.1, vjust=0) + 
  scale_color_manual(values = palette) + theme_light()

b <- integrationMetricsSummary %>% dplyr::filter(Metric %in% c("norm_cLISI","CiLISI")) %>%
  pivot_wider(names_from = Metric, values_from = value ) %>%
  ggplot(aes(x=CiLISI, y=`norm_cLISI`, label=Method)) +
  geom_point(aes(color=Method), size=2) + geom_text(hjust=0.1, vjust=0) + 
  scale_color_manual(values = palette) + theme_light()

a | b

ggsave("plots/DC_metrics_scatterplot_expertAnnotation.pdf", height=3, width=10)
```