
# Install dependencies if needed
Let's use latest version of Seurat and ProjecT

## General packages
```{r}
gen.pck <- c("Seurat", "remotes", "BiocManager",
             "dplyr", "tidyr", "tibble", "ggplot2")

sapply(gen.pck,
       function(p){
         if(!require(p, character.only = T)){
           # use renv to install to keep of record of packages
           # moreover, renv allow to install from both CRAN and bioconductor
           renv::install(p)
         }
         library(p,
                 character.only = T)
       })

```

## Carmona lab packages
Let's install dev versions for some of them, as they have some modifications we need.
You will be prompted whether to upgrade other packages, up to you.

```{r}
# master version packages
csi.pck <- c("SignatuR", "UCell", "STACAS", "scGate")

sapply(csi.pck,
       function(p){
         if(!require(p, character.only = T)){
           # install from github
           remotes::install_github(paste0("carmonalab/", p))
         }
         library(p,
                 character.only = T)
       })

csi.pck <- c("ProjecTILs", "HiTME")

sapply(csi.pck,
       function(p){
         if(!require(p, character.only = T)){
           # install from github, dev version
           remotes::install_github(paste0("carmonalab/", p),
                                   ref = "dev")
         }
         library(p,
                 character.only = T)
       })
```

# Load previous DC reference map
```{r}
path_ref <- "~/Dropbox/CSI/reference_atlases" 
DC = ProjecTILs::load.reference.map(file.path(path_ref, "DC_human_ref_v1.rds"))
```

```{r}
up <- UMAPPlot(DC,
               cols = DC@misc$atlas.palette) + NoAxes()
up
```


Count of each celltype
```{r}
p1 <- DC@meta.data %>% 
        count(functional.cluster) %>% 
        mutate(freq = n/sum(n)) %>% 
        ggplot(aes(functional.cluster, freq,
                   fill = functional.cluster)) +
        ylim(c(0,0.6)) +
        geom_col(show.legend = F) +
        labs(y = "Relative proportions of DC subtypes") +
        scale_fill_manual(values = DC@misc$atlas.palette) +
        theme_bw()

all <- DC@meta.data %>% 
        count(functional.cluster, Study) %>% 
        group_by(Study) %>%
        mutate(freq = n/sum(n)) %>%
        ggplot(aes(Study, freq, fill = functional.cluster)) +
        geom_col() +
        scale_fill_manual(values = DC@misc$atlas.palette) +
        labs(y = "Proportion of cells") +
        theme_bw()

allp <- DC@meta.data %>% 
        count(functional.cluster, Study) %>% 
        group_by(functional.cluster) %>%
        mutate(freq = n/sum(n)) %>%
        ggplot(aes(functional.cluster, freq, fill = Study)) +
        geom_col() +
        labs(y = "Proportion of cells") +
        theme_bw()

```

# Remove half of cDC2_CLEC10A
Remove randomly half of this cell type
```{r}
ctype <- "cDC2_CLEC10A"
# vector of all cells
allcl <- Cells(DC)

# vector of cDC2_CLEC1A
cl <- Cells(DC[, DC$functional.cluster == ctype])

#fix seed
set.seed(22)

# randomly peak of these cell types that will be removed
cl_subset <- sample(cl,
                    size = ceiling(length(cl)/1.5),
                    replace = F)
keep <- allcl[!allcl %in% cl_subset]
DC2 <- subset(DC, cells = keep)
```

## UMAPS
```{r}
up2 <- UMAPPlot(DC2) + NoAxes()
up2
```

### Recalculate refmap
```{r, eval = F}
# keep parameters of old version
DC2 <- make.reference(DC2,
                      assay = "integrated",
                      atlas.name = "DC_human_v2",
                      ndim = 40,
                      nfeatures = 2000,
                      recalculate.umap = T,
                      scGate.model.human = DC@misc$scGate$human,
                      scGate.model.mouse = DC@misc$scGate$mouse,
                      layer1_link = "CL:0000451")
```

```{r}
up3 <- UMAPPlot(DC2) + NoAxes()
```

Count of each celltype
```{r}
p1 <- DC2@meta.data %>% 
        count(functional.cluster) %>% 
        mutate(freq = n/sum(n)) %>% 
        ggplot(aes(functional.cluster, freq,
                   fill = functional.cluster)) +
        geom_col(show.legend = F) +
        ylim(c(0,0.6)) +
        labs(y = "Relative proportions of DC subtypes") +
        scale_fill_manual(values = DC@misc$atlas.palette) +
        theme_bw()

all <- DC2@meta.data %>% 
        count(functional.cluster, Study) %>% 
        group_by(Study) %>%
        mutate(freq = n/sum(n)) %>%
        ggplot(aes(Study, n, fill = functional.cluster)) +
        geom_col() +
        scale_fill_manual(values = DC@misc$atlas.palette) +
        labs(y = "Number of cells") +
        theme_bw()

allp <- DC2@meta.data %>% 
        count(functional.cluster, Study) %>% 
        group_by(functional.cluster) %>%
        mutate(freq = n/sum(n)) %>%
        ggplot(aes(functional.cluster, freq, fill = Study)) +
        geom_col() +
        labs(y = "Proportion of cells") +
        theme_bw()

```

# Split MonoDC
```{r}
monoDC <- DC[,DC$functional.cluster == "MonoDC"]

UMAPPlot(monoDC)
```

## Recompute space

```{r}
var.features = 2000
ndim = 40
vars.regress = c("percent.mito", "nFeature_RNA")

monoDC <- monoDC %>% 
        # FindVariableFeatures(selection.method = "vst",
        #                      nfeatures = var.features,
        #                      verbose = F) %>% 
        # ScaleData(vars.to.regress = vars.regress,
        #           assay = "integrated") %>% 
        RunPCA(npcs=ndim,
               verbose = F) %>% 
        RunUMAP(reduction = "pca",
                dims = 1:ndim,
                verbose = F)

upp <- UMAPPlot(monoDC)
upp
```

## Cluster
```{r}
monoDC <- FindNeighbors(monoDC,
                        dims = 1:20)
monoDC <- FindClusters(monoDC,
                       resolution = 0.2)
UMAPPlot(monoDC) 

monoDC$clus <- factor(monoDC$seurat_clusters,
                      levels = c(0,1, 2),
                      labels = c("MonoDC_CD14",
                                 "MonoDC_DN",
                                 "MonoDC_CD16"))

(upcl <- UMAPPlot(monoDC,
                  group.by = "clus") +
        ggtitle("MonoDC"))
```

## Feature plot
```{r, fig.height=10}
feats <- c("CD14", "CD1C", "THBD", "FCGR3A",
           "SERPINA1", "FTL", "IFITM3", "CSF1R",
           "CLEC10A", "FCER1A")
DefaultAssay(monoDC) <- "RNA"
fpl <- FeaturePlot(monoDC,
                   features = feats)
fpl
```

```{r, fig.width=4, fig.height=7}
vln <- VlnPlot(monoDC,
               features = feats,
               group.by = "clus",
               stack = T, flip = T) +
        NoLegend()
vln
```







# Rename cell types
```{r}
MonoDC_CD14 <- Cells(monoDC[, monoDC$clus == "MonoDC_CD14"])
MonoDC_CD16 <- Cells(monoDC[, monoDC$clus == "MonoDC_CD16"])
MonoDC_DN <- Cells(monoDC[, monoDC$clus == "MonoDC_DN"])

DC2@meta.data <- DC2@meta.data %>% 
        rownames_to_column("id") %>% 
        rowwise() %>% 
        mutate(functional.cluster = as.character(functional.cluster),
               functional.cluster = ifelse(id %in% MonoDC_CD14,
                                           "MonoDC_CD14",
                                           ifelse(id %in% MonoDC_CD16,
                                                  "MonoDC_CD16",
                                                  ifelse(id %in% MonoDC_DN,
                                                         "MonoDC_DN",
                                                         ifelse(functional.cluster == "DC3",
                                                                "cDC_LAMP3_CCR7",
                                                                functional.cluster
                                                         )
                                                  )
                                           )
               ),
               functional.cluster = factor(functional.cluster,
                                           levels = c("cDC1",
                                                      "cDC2_CLEC10A",
                                                      "cDC2_CD1A",
                                                      "cDC_LAMP3_CCR7",
                                                      "pDC",
                                                      "AS-DC",
                                                      "MonoDC_CD14",
                                                      "MonoDC_CD16",
                                                      "MonoDC_DN"))
        ) %>% 
        ungroup() %>% 
        column_to_rownames("id")    
```

## Make reference map
Color palette
```{r}
require(RColorBrewer)
palette <- brewer.pal(n=length(unique(DC2$functional.cluster)), name="Set2")

names(palette) <- levels(DC2$functional.cluster)
```


```{r}
         
# keep parameters of old version
DC2 <- make.reference(DC2,
                      assay = "integrated",
                      atlas.name = "DC_human_v2",
                      ndim = 40,
                      nfeatures = 2000,
                      recalculate.umap = T,
                      scGate.model.human = DC@misc$scGate$human,
                      scGate.model.mouse = DC@misc$scGate$mouse,
                      layer1_link = "CL:0000451",
                      color.palette = palette)
```

```{r}
upfi <- UMAPPlot(DC2,
                 cols = palette) +
        NoAxes()
upfi
```

Count of each celltype
```{r}
p1 <- DC2@meta.data %>% 
        count(functional.cluster) %>% 
        mutate(freq = n/sum(n)) %>% 
        ggplot(aes(functional.cluster, freq,
                   fill = functional.cluster)) +
        geom_col(show.legend = F,
                 color = "black") +
        scale_fill_manual(values = palette) +
        labs(y = "Relative proportions of DC subtypes") +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45,
                                         hjust = 1, 
                                         vjust = 1))

```

Check even distribution of studies
```{r}
all <- DC2@meta.data %>% 
        count(functional.cluster, Study) %>% 
        group_by(Study) %>%
        mutate(freq = n/sum(n)) %>%
        ggplot(aes(Study, freq, fill = functional.cluster)) +
        geom_col(color = "black") +
        scale_fill_manual(values = DC2@misc$atlas.palette) +
        labs(y = "Proportion of cells") +
        theme_bw()

allp <- DC2@meta.data %>% 
        count(functional.cluster, Study) %>% 
        group_by(functional.cluster) %>%
        mutate(freq = n/sum(n)) %>%
        ggplot(aes(functional.cluster, freq, fill = Study)) +
        geom_col(color = "black") +
        labs(y = "Proportion of cells") +
        theme_bw()
```

### Dotplot
```{r}
v2 <- DotPlot(lir$DCv2,
              assay = "RNA",
              cols = c("white", "red"),
              group.by = "functional.cluster",
              features = unique(genes4radar)) +
        coord_flip() +
        theme(axis.text.x = element_text(angle = 45,
                                         hjust = 1,
                                         vjust = 1))

```


## save reference
```{r}
fpath <- "~/Dropbox/CSI/reference_atlases/DC_human_ref_v2.rds"
saveRDS(DC2, fpath)
```


# Test on Brown DC dataset
## Load new reference
```{r}
DC2 <- ProjecTILs::load.reference.map(fpath)
```

## Load Brown
```{r}
ddir <- "~/Documents/Projects/Supercell_maps/HiTME/data/Brown2019_humanDC_data/Brown2019.rds"
brown.seurat <- readRDS(ddir)
```

## annotation
```{r}
lir <- list("DCv1" = DC,
            "DCv2" = DC2)

rm(DC, DC2);gc()

for(r in names(lir)){
        brown.seurat <- ProjecTILs.classifier(brown.seurat,
                                              ref = lir[[r]],
                                              filter.cells = F)
        
        brown.seurat@meta.data[[r]] <- brown.seurat@meta.data$functional.cluster
        brown.seurat@meta.data[[paste0(r, ".conf")]] <- brown.seurat@meta.data$functional.cluster.conf
}
```

## Evaluation
```{r}
brown.seurat <- NormalizeData(brown.seurat) |>
  ScaleData() |>
  FindVariableFeatures() |>
  RunPCA(npcs = 20) |>
  RunUMAP(dims = 1:20)
```

### UMAPs
```{r}
ups <- UMAPPlot(brown.seurat,
                group.by = c("cell_type",
                             "DCv1","DCv2"))
ups
```
### Radar plots
```{r}
genes4radar <- c("CLEC9A","XCR1","CADM1","CLEC10A","CD1C","CD1A",
                 "LAMP3","CCR7","FSCN1","S100A8","S100A9","VCAN","FCN1",
                 "GZMB","TSPAN13","TCF4","PPP1R14A","AXL", feats) 

rad.li <- list()
for(r in names(lir)){
        rad.li[[r]] <- plot.states.radar(lir[[r]],
                                         query=brown.seurat,
                                         genes4radar = genes4radar,
                                         min.cells = 5)
}
```


### Violin plots
```{r}
vl.list <- list()

for(r in names(lir)){
        vl.list[[r]] <- VlnPlot(brown.seurat,
                                group.by = r,
                                features = genes4radar,
                                stack = T,
                                flip = T) +
                NoLegend() +
                labs(x = "") +
                scale_x_discrete(limits = levels(lir[[r]]$functional.cluster)) +
                ggtitle(r)
        
}
vl.list$DCv1 + vl.list$DCv2

```





### Confusion matrix
```{r}
plot_accuracy <- function(data,
                          ground_truth_var,
                          prediction_var,
                          relative = T,
                          useNA = "no",
                          xlab = NULL,
                          ylab = NULL,
                          plot.title = "") {
  if(is.null(xlab)){
    xlab <-  ground_truth_var
  }
  if(is.null(ylab)){
    ylab <-  prediction_var
  }
  legend.title <- "Number of cells"
  
  data <- as.data.frame(data)
  # Get counts for every group
  pz <- table(data[,ground_truth_var],
              data[,prediction_var],
              useNA = useNA)
  
  if(relative){
    pz <- prop.table(pz, margin = 1) %>% 
      round(digits =2)
    legend.title <- "Proportion of cells"
  }
  
  cells <- pz %>% as.data.frame() %>% 
    ggplot(aes(Var1, Var2,
               fill = Freq)) + 
    geom_tile() +
    scale_fill_gradient(name = legend.title,
                        low = "#ffffc8",
                        high = "#7d0025") +
    labs(x = xlab,
         y = ylab) +
    geom_label(aes(label = Freq),
               color = "white", fill = "grey15") +
    ggtitle(plot.title) +
    cowplot::theme_cowplot() +
    theme(axis.text.x = element_text(angle = 45,
                                     hjust = 1,
                                     vjust = 1))
  
  return(cells)
  
}
```

```{r}
comps <- c("cell_type", "DCv1", "DCv2")
d <- c()

pa.list <- list()

for(a in comps){
        d <- c(d, a)
        for(b in comps[!comps %in% d]){
             pa.list[[paste(a,b, sep = "_vs_")]] <- 
                     plot_accuracy(brown.seurat@meta.data,
                                   a,
                                   b,
                                   relative = T)
        }
}

cpa <- ggpubr::ggarrange(plotlist = pa.list,
                         common.legend = T,
                         legend = "bottom",
                         ncol = 3)
```

# pseudobulk
## Get pseudobulk
```{r}
ps <- AverageExpression(lir$DCv2,
                          assays = "RNA",
                          group.by = c("Study","functional.cluster")
                          )[["RNA"]]
```


## Desnormalize counts
```{r, eval = F}
source("../scRNAseq_data_processing/Rutils/ReverseNormalize.R")

raw <- apply(DC@assays$RNA$counts, 2, reverseNormalize)

raw[1:50,1:10] %>% View()
DC@assays$RNA$counts[1:50,1:10] %>% as.matrix() %>% View()


DC.raw <- CreateSeuratObject(counts = raw,
                             meta.data = DC@meta.data)

ps <- AggregateExpression(DC.raw,
                              group.by = c("Study","functional.cluster")
                          )[["RNA"]]
```

!! Normalization reverse not possible, as some datasets are from SmartSeq2 technology

```{r}
metadata <- data.frame(rn = colnames(ps)) %>%
        tidyr::separate(rn, sep = "_",
                        into = c("Study", "celltype"),
                        remove = F) %>%
        tibble::column_to_rownames("rn")
```




## Get PCA
Normalize
```{r}

nVarGenes = 500

# remove black listed genes
load("../HiTME/data/default_black_list.RData")
ps <- ps[!rownames(ps) %in% c(unlist(black.list), "MSB4X", "TMSB4X", "B2M"),]

# select most variable genes
rv <- MatrixGenerics::rowVars(ps)
select <- order(rv, decreasing=TRUE)[seq_len(min(nVarGenes, length(rv)))]
select <- rownames(ps)[select]

vsd <- ps[select[select %in% rownames(ps)],]

pc <- stats::prcomp(t(vsd))

fviz_pca(pc)
```



Do PCA
```{r}
# plot for PCA
pc_sum <- summary(pc)
PC1_varexpl <- pc_sum$importance[2,"PC1"]
PC2_varexpl <- pc_sum$importance[2,"PC2"]

# get first 2 PC
pc.df <- pc$x[,1:2] %>% as.data.frame() %>%
        tibble::rownames_to_column("sample_celltype") %>%
        left_join(., metadata %>% tibble::rownames_to_column("sample_celltype"),
                  by = "sample_celltype")

pc.pl <- pc.df %>%
        ggplot2::ggplot(ggplot2::aes(PC1, PC2,
                                     color = Study)) +
        ggplot2::geom_point() +
        ggplot2::guides(color=guide_legend(ncol=2))+
        labs(
             y = paste0("PC2 (", PC2_varexpl*100, " %)"),
             x = paste0("PC1 (", PC1_varexpl*100, " %)"))+
        ggplot2::theme(
                panel.background = element_rect(fill = "white"),
                legend.key = element_rect(fill = "white")
        )
pc.pl
```

```{r}
up <- umap::umap(pc$x[,1:10])

up.df <- up$layout
colnames(up.df) <- c("UMAP_1", "UMAP_2")

up.df <- up.df %>% 
        as.data.frame() %>% 
        tibble::rownames_to_column("sample_celltype") %>%
        left_join(., metadata %>% tibble::rownames_to_column("sample_celltype"),
                  by = "sample_celltype")

up.pl <- up.df %>%
        ggplot2::ggplot(ggplot2::aes(UMAP_1, UMAP_2,
                                     color = Study)) +
        ggplot2::geom_point() +
        ggplot2::guides(color=guide_legend(ncol=2))+
        ggplot2::theme(
                panel.background = element_rect(fill = "white"),
                legend.key = element_rect(fill = "white")
        )
up.pl
```

## Sex of the samples
```{r}
sgenes <- c("XIST", "RPS4Y1")
ps.sex <- AverageExpression(DC,
                            group.by = "Study")[["RNA"]][sgenes,] %>% 
  t() %>% 
  as.data.frame()


ps.sex <- ps.sex %>% 
  mutate(sex = log2(XIST/RPS4Y1))
```

## Heatmap pseudobulk
Functions below not working, is it for the Seurat version??
```{r, eval = F}

remotes::install_version("Seurat", version = "4.3.0.1")
remotes::install_version("SeuratObject", version = "4.1.3")
.rs.restartR()
```



```{r}
require(ProjecTILs)
require(BiocParallel)

# top most DE genes
top <- 10


heatli <- bplapply(names(lir),
                   BPPARAM = MulticoreParam(workers = 2, progressbar = T),
                   function(r){
                     DefaultAssay(lir[[r]]) <- "RNA"
                     
                     gmark <- NormalizeData(lir[[r]]) %>% 
                       FindAllMarkers.bygroup(
                         assay = "RNA",
                         split.by = "Study",
                         min.cells.group = 5,
                         min.pct = 0.5,
                         min.diff.pct = 0.2,
                         logfc.threshold = 1,
                         min.freq = 0.4)
                     
                     gene.use <- lapply(gmark,
                                        function(df){
                                          df %>% 
                                            arrange(desc(avg.FC)) %>% 
                                            head(top) %>% 
                                            pull(Gene)
                                        }) %>%
                       unlist() 
                     
                     h <- celltype.heatmap(lir[[r]],
                                           group.by = "Study",
                                           genes = gene.use,
                                           min.cells = 5,
                                          cluster_samples = T)
                     
                     return(h)
                     
                   })
names(heatli) <- names(lir)

```


```{r}
tt = FindAllMarkers(lir$DCv1)
```

















