
# Install dependencies if needed
Let's use latest version of Seurat and ProjecT

## General packages
```{r}
gen.pck <- c("Seurat", "remotes", "BiocManager",
             "dplyr", "tidyr", "tibble", "ggplot2")

sapply(gen.pck,
       function(p){
         if(!require(p, character.only = T)){
           # use renv to install to keep of record of packages
           # moreover, renv allow to install from both CRAN and bioconductor
           renv::install(p)
         }
         library(p,
                 character.only = T)
       })

```

## Carmona lab packages
Let's install dev versions for some of them, as they have some modifications we need.
You will be prompted whether to upgrade other packages, up to you.

```{r}
# master version packages
csi.pck <- c("SignatuR", "UCell", "STACAS", "scGate")

sapply(csi.pck,
       function(p){
         if(!require(p, character.only = T)){
           # install from github
           remotes::install_github(paste0("carmonalab/", p))
         }
         library(p,
                 character.only = T)
       })

csi.pck <- c("ProjecTILs")

sapply(csi.pck,
       function(p){
         if(!require(p, character.only = T)){
           # install from github, dev version
           remotes::install_github(paste0("carmonalab/", p),
                                   ref = "dev")
         }
         library(p,
                 character.only = T)
       })
```

# Load previous DC reference map
```{r}
path_ref <- "~/Dropbox/CSI/reference_atlases" 
DC = ProjecTILs::load.reference.map(file.path(path_ref, "DC_human_ref_v1.rds"))
```

```{r}
up <- UMAPPlot(DC) + NoAxes()
```


Count of each celltype
```{r}
p1 <- DC@meta.data %>% 
        count(functional.cluster) %>% 
        mutate(freq = n/sum(n)) %>% 
        ggplot(aes(functional.cluster, freq,
                   fill = functional.cluster)) +
        ylim(c(0,0.6)) +
        geom_col(show.legend = F) +
        labs(y = "Relative proportions of DC subtypes") +
        scale_fill_manual(values = DC@misc$atlas.palette) +
        theme_bw()

all <- DC@meta.data %>% 
        count(functional.cluster, Study) %>% 
        group_by(Study) %>%
        mutate(freq = n/sum(n)) %>%
        ggplot(aes(Study, freq, fill = functional.cluster)) +
        geom_col() +
        scale_fill_manual(values = DC@misc$atlas.palette) +
        labs(y = "Proportion of cells") +
        theme_bw()

allp <- DC@meta.data %>% 
        count(functional.cluster, Study) %>% 
        group_by(functional.cluster) %>%
        mutate(freq = n/sum(n)) %>%
        ggplot(aes(functional.cluster, freq, fill = Study)) +
        geom_col() +
        labs(y = "Proportion of cells") +
        theme_bw()

```

# Remove half of cDC2_CLEC10A
Remove randomly half of this cell type
```{r}
ctype <- "cDC2_CLEC10A"
# vector of all cells
allcl <- Cells(DC)

# vector of cDC2_CLEC1A
cl <- Cells(DC[, DC$functional.cluster == ctype])

#fix seed
set.seed(22)

# randomly peak of these cell types that will be removed
cl_subset <- sample(cl,
                    size = ceiling(length(cl)/1.5),
                    replace = F)
keep <- allcl[!allcl %in% cl_subset]
DC2 <- subset(DC, cells = keep)
```

## UMAPS
```{r}
up2 <- UMAPPlot(DC2) + NoAxes()
```

### Recalculate refmap
```{r}
# keep parameters of old version
DC2 <- make.reference(DC2,
                      assay = "integrated",
                      atlas.name = "DC_human_v2",
                      ndim = 40,
                      nfeatures = 2000,
                      recalculate.umap = T,
                      scGate.model.human = DC@misc$scGate$human,
                      scGate.model.mouse = DC@misc$scGate$mouse,
                      layer1_link = "CL:0000451")
```

```{r}
up3 <- UMAPPlot(DC2) + NoAxes()
```

Count of each celltype
```{r}
p1 <- DC2@meta.data %>% 
        count(functional.cluster) %>% 
        mutate(freq = n/sum(n)) %>% 
        ggplot(aes(functional.cluster, freq,
                   fill = functional.cluster)) +
        geom_col(show.legend = F) +
        ylim(c(0,0.6)) +
        labs(y = "Relative proportions of DC subtypes") +
        scale_fill_manual(values = DC@misc$atlas.palette) +
        theme_bw()

all <- DC2@meta.data %>% 
        count(functional.cluster, Study) %>% 
        group_by(Study) %>%
        mutate(freq = n/sum(n)) %>%
        ggplot(aes(Study, n, fill = functional.cluster)) +
        geom_col() +
        scale_fill_manual(values = DC@misc$atlas.palette) +
        labs(y = "Number of cells") +
        theme_bw()

allp <- DC2@meta.data %>% 
        count(functional.cluster, Study) %>% 
        group_by(functional.cluster) %>%
        mutate(freq = n/sum(n)) %>%
        ggplot(aes(functional.cluster, freq, fill = Study)) +
        geom_col() +
        labs(y = "Proportion of cells") +
        theme_bw()

```

# Split MonoDC
```{r}
monoDC <- DC[,DC$functional.cluster == "MonoDC"]

UMAPPlot(monoDC)
```

## Recompute space

```{r}
var.features = 2000
ndim = 40
vars.regress = c("percent.mito", "nFeature_RNA")

monoDC <- monoDC %>% 
        # FindVariableFeatures(selection.method = "vst",
        #                      nfeatures = var.features,
        #                      verbose = F) %>% 
        # ScaleData(vars.to.regress = vars.regress,
        #           assay = "integrated") %>% 
        RunPCA(npcs=ndim,
               verbose = F) %>% 
        RunUMAP(reduction = "pca",
                dims = 1:ndim,
                verbose = F)

upp <- UMAPPlot(monoDC)
upp
```

## Cluster
```{r}
monoDC <- FindNeighbors(monoDC,
                        dims = 1:20)
monoDC <- FindClusters(monoDC,
                       resolution = 0.1)
monoDC$clus <- factor(monoDC$seurat_clusters,
                      levels = c(0,1),
                      labels = c("MonoDC_CD14", "MonoDC_CD16"))

(upcl <- UMAPPlot(monoDC,
                  group.by = "clus") +
        ggtitle("MonoDC"))
```

## Feature plot
```{r}
feats <- c("CD14", "CD1C", "THBD", "FCGR3A",
           "SERPINA1", "FTL", "IFITM3", "CSF1R",
           "CLEC10A", "FCER1A")
DefaultAssay(monoDC) <- "RNA"
fpl <- FeaturePlot(monoDC,
                   features = feats)
fpl
```

```{r}
vln <- VlnPlot(monoDC,
               features = feats,
               group.by = "clus",
               stack = T, flip = T) +
        NoLegend()
```







# Rename cell types
```{r}
MonoDC_CD14 <- Cells(monoDC[, monoDC$clus == "MonoDC_CD14"])
MonoDC_CD16 <- Cells(monoDC[, monoDC$clus == "MonoDC_CD16"])

DC2@meta.data <- DC2@meta.data %>% 
        rownames_to_column("id") %>% 
        rowwise() %>% 
        mutate(functional.cluster = as.character(functional.cluster),
               functional.cluster = ifelse(id %in% MonoDC_CD14,
                                           "MonoDC_CD14",
                                           ifelse(id %in% MonoDC_CD16,
                                           "MonoDC_CD16",
                                           ifelse(functional.cluster == "DC3",
                                                  "cDC_LAMP3_CCR7",
                                                  functional.cluster
                                                  )
                                           )
                                           )
               ) %>% 
        ungroup() %>% 
        column_to_rownames("id")    
```

## Make reference map
```{r}
pal <- c(DC@misc$atlas.palette[-7],
         "MonoDC_CD14" = "#E31A1C",
         
         "#1F78B4"
         
# keep parameters of old version
DC2 <- make.reference(DC2,
                      assay = "integrated",
                      atlas.name = "DC_human_v2",
                      ndim = 40,
                      nfeatures = 2000,
                      recalculate.umap = T,
                      scGate.model.human = DC@misc$scGate$human,
                      scGate.model.mouse = DC@misc$scGate$mouse,
                      layer1_link = "CL:0000451")
```

```{r}
upfi <- UMAPPlot(DC2)
upfi
```

Count of each celltype
```{r}
p1 <- DC2@meta.data %>% 
        count(functional.cluster) %>% 
        mutate(freq = n/sum(n)) %>% 
        ggplot(aes(functional.cluster, freq,
                   fill = functional.cluster)) +
        geom_col(show.legend = F) +
        labs(y = "Relative proportions of DC subtypes") +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45,
                                         hjust = 1, 
                                         vjust = 1))

```



