---
title: "Load all datasets in Gerhard et al. metanalysis, and convert to Seurat"
author: "Massimo Andreatta & Santiago Carmona"
#output: html_notebook
---

```{r}
renv::restore()
```

Libraries
```{r}
#remotes::install_github('carmonalab/UCell')
#remotes::install_github('carmonalab/scGate')

library(ggplot2)
library(Seurat)
library(tidyr)
library(SeuratObject)
library(Matrix)
library(patchwork)
library(remotes)
library(UCell)
library(scGate)
```

#NB: for conversion between Scanpy format to matrix files, see export_data.py script

Read in data for individual studies
```{r}
pdir <- "~/Dropbox/CSI/Datasets/Gerhard_JEM"
ddir <- "~/Dropbox/CSI/Datasets/Gerhard_JEM/exported"

studies <- list.dirs(path=ddir, full.names = FALSE)
studies <- studies[nchar(studies)>0]

names <- gsub("_\\S+","",studies, perl=T)

seurat.list <- list()

for (i in seq_along(studies)) {
  s = studies[i]
  proj = names[i]
   
  data <- t(readMM(file = sprintf("%s/%s/matrix.mtx", ddir, s)))

  cell.names <- read.table(sprintf("%s/%s/barcodes.tsv", ddir, s), sep="\t", header=T)[,1]
  gene.names <- read.table(sprintf("%s/%s/genes.tsv", ddir, s), sep="\t", header=T)[,1]
  meta <- read.table(sprintf("%s/%s/metadata.csv",ddir,s), sep=",",header=T)[,-1]
  rownames(meta) <- cell.names
  rownames(data) <- gene.names
  colnames(data) <- cell.names 
   
  seurat.list[[proj]] <- CreateSeuratObject(counts=data, project=proj, assay="RNA", meta.data = meta)
  seurat.list[[proj]]$Study <- proj
  
  #mito and ribo content
  patterns <- c("^RP[LS]","^MT-")

  pc.ribo <- PercentageFeatureSet(seurat.list[[proj]], pattern = patterns[1])
  pc.mito <- PercentageFeatureSet(seurat.list[[proj]], pattern = patterns[2])
  
  seurat.list[[proj]] <- AddMetaData(seurat.list[[proj]], metadata = pc.ribo, col.name = "percent.ribo")
  seurat.list[[proj]] <- AddMetaData(seurat.list[[proj]], metadata = pc.mito, col.name = "percent.mito")
  
}

##And save the list of object to disk

#saveRDS(seurat.list, file="Gerhard_metanalysis_seuratlist.rds")
saveRDS(seurat.list, file=sprintf("%s/Gerhard_metanalysis_seuratlist.rds",pdir))
```

Data in counts slot appear to be already normalized?

How is that possible, if 10x files were read in?
```{r}
lapply(seurat.list, function(x) {
  c <- colSums(x@assays$RNA@counts)
  head(c)
})


#maybe normalized, but not log-normalized (?)
hist(seurat.list[[1]]@assays$RNA@counts["CD74",], breaks=50)
hist(seurat.list[[1]]@assays$RNA@counts["CCR7",], breaks=50)
hist(seurat.list[[1]]@assays$RNA@counts["XCR1",], breaks=50)
```



Basic exploration of data
```{r}
plots <- list()

ndim=30
for (i in seq_along(seurat.list)) {
  s = names(seurat.list)[i]
  
  obj <- seurat.list[[s]]
  obj <- NormalizeData(obj) #Counts are not integers...? already normalized?
  
  obj <- FindVariableFeatures(obj, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
  obj <- ScaleData(obj, do.scale = TRUE, do.center = TRUE, verbose = F) 
  obj <- RunPCA(object = obj, features = obj@assays$RNA@var.features, npcs = ndim, verbose = F)
  
  obj <- RunUMAP(obj, reduction = "pca", dims = 1:ndim, seed.use=123, verbose = F)
  plots[[s]] <- DimPlot(obj, reduction = "umap", group.by = "state", label = T) + ggtitle(s) + NoLegend() +
             theme(axis.title=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
}
```

```{r fig.height=8, fig.width=16}
wrap_plots(plots, ncol=4)

ggsave("plots/Gerhard_sets_unsup_umap.pdf", width = 24, height=15)
```


#Focus on DCs only
```{r}
lapply(seurat.list, function(x) {unique(x$state)})

# brownSpleen, martin, qian, qianbc, qiancrc, qianlung, qianovarian are focused on DCs only

# for the rest, subset on DCs

dc.seurat <- seurat.list

dc.seurat$villani <- subset(dc.seurat$villani, subset=state %in% c("vi-DC1",
                                                                   "vi-DC5",
                                                                   "vi-DC2",
                                                                   "vi-DC3",
                                                                   "vi-DC6",
                                                                   "vi-DC4"))
dc.seurat$zhang <- subset(dc.seurat$zhang, subset=state %in% c("zhH-DC-C2-FCER1A",
                                                               "zhH-DC-C4-LAMP3",
                                                               "zhH-DC-C3-CLEC9A",
                                                               "zhH-DC-C1-CD1C" ))

dc.seurat$zhangcrc <- subset(dc.seurat$zhangcrc, subset=state %in% c("zhC-hM03_cDC2-CD1C",
                                                                     "zhC-hM04_cDC1-BATF3",
                                                                     "zhC-hM02_pDC-LILRA4" ))

dc.seurat$zilionis <- subset(dc.seurat$zilionis, subset=state %in% c("zi-tMonoDC",
                                                                     "zi-tDC1",
                                                                     "zi-tDC2",
                                                                     "zi-tDC3",
                                                                     "zi-tpDC"))

saveRDS(dc.seurat, file=sprintf("%s/Gerhard_metanalysis_seuratlist.DC.rds",pdir))
```

Basic exploration of data
```{r}
plots <- list()

ndim=15
for (s in names(dc.seurat)) {
  
  dc.seurat[[s]] <- NormalizeData(dc.seurat[[s]])
  dc.seurat[[s]] <- FindVariableFeatures(dc.seurat[[s]], selection.method = "vst", nfeatures = 500, verbose = FALSE)
  dc.seurat[[s]] <- ScaleData(dc.seurat[[s]], verbose = F) |> RunPCA(npcs = ndim, verbose=F)
  
  dc.seurat[[s]] <- RunUMAP(dc.seurat[[s]], reduction = "pca", dims = 1:ndim, seed.use=123, verbose = F)
  plots[[s]] <- DimPlot(dc.seurat[[s]], reduction = "umap", group.by = "state", label = T) + ggtitle(s) + NoLegend() +
             theme(axis.title=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
}
```

```{r fig.height=8, fig.width=16}
wrap_plots(plots, ncol=4)

ggsave("plots/Gerhard_sets_unsup_DC.pdf", width = 24, height=15)
```


```{r fig.height=3}
table(dc.seurat$zilionis$Patient)
a <- DimPlot(dc.seurat$zilionis, reduction="umap", group.by = "Patient") + theme(aspect.ratio = 1)
b <- DimPlot(dc.seurat$zilionis, reduction="umap", group.by = "state") + theme(aspect.ratio = 1)

a | b
```

```{r}
dset <- "brownSpleen"
#cDC1
FeaturePlot(dc.seurat[[dset]], features=c("CLEC9A","XCR1","BATF3","C1orf54"))
FeaturePlot(dc.seurat[[dset]], features=c("CPNE3","CPVL","CADM1","WDFY4"))

#cDC2
FeaturePlot(dc.seurat[[dset]], features=c("CD1C","SIRPA","ITGAM","CD207"))
FeaturePlot(dc.seurat[[dset]], features=c("CLEC10A","AIF1","FCER1A","CD1A"))
#DC3
FeaturePlot(dc.seurat[[dset]], features=c("CCR7","LAMP3","FSCN1","BIRC3"))
#pDC
FeaturePlot(dc.seurat[[dset]], features=c("IRF7","PLAC8","LILRA4","TCF4"))
FeaturePlot(dc.seurat[[dset]], features=c("IRF7","GZMB","IL3RA","IRF8"))

FeaturePlot(dc.seurat[[dset]], features=c("nFeature_RNA","nCount_RNA","percent.ribo","percent.mito"))

```

Can we build simple signatures based on marker genes?
```{r}
sigs <- list()

sigs$cDC1 <- c("CLEC9A","XCR1","BATF3","CADM1")
sigs$cDC2 <- c("CD207","CD1C","CLEC10A","FCER1A")
sigs$DC3 <- c("LAMP3","CCR7","FSCN1","IL3RA")
sigs$pDC <- c("IRF7","LILRA4","TCF4")

dc.seurat <- lapply(dc.seurat, function(x) {
  dc.seurat <- AddModuleScore_UCell(x, features=sigs, name=NULL)
  SmoothKNN(dc.seurat, signature.names = names(sigs), reduction = "pca")
})
```

```{r fig.height=3}
for (s in names(dc.seurat)) {
  p <- FeaturePlot(dc.seurat[[s]], features=paste0(names(sigs),"_kNN"))
  plot(p)
}
```
Try assigning labels based on scGate (simple models)
```{r}
ss <- list()
ss$cDC1 <- gating_model(name="cDC1", signature = sigs$cDC1)
ss$cDC2 <- gating_model(name="cDC2", signature = sigs$cDC2)
ss$DC3 <- gating_model(name="DC3", signature = sigs$DC3)
ss$pDC <- gating_model(name="pDC", signature = sigs$pDC)

dc.seurat <- lapply(dc.seurat, function(x) {
  scGate(x, model=ss, reduction = "pca", ncores = 4)
})
```


```{r fig.height=12, fig.width=4}
library(RColorBrewer)
palette <- brewer.pal(n=5, name="Set3")
names(palette) <- c(names(sigs), "Multi")

pll <- list()
i <- 0
for (s in names(dc.seurat)) {
  pll[[i <- i+1]] <- DimPlot(dc.seurat[[s]], group.by = "scGate_multi", cols = palette) + theme(aspect.ratio = 1)
  pll[[i <- i+1]] <- DimPlot(dc.seurat[[s]], group.by = "state") + theme(aspect.ratio = 1) + ggtitle(s)
}
wrap_plots(pll, ncol = 2)

ggsave("plots/Gerhard_sets_scGate_multi.pdf", height=32, width=11)
```

Save - with annotations
```{r}
saveRDS(dc.seurat, file=sprintf("%s/Gerhard_metanalysis_seuratlist.DC.anno.rds",pdir))
```
